Learning from Massive Noisy Labeled Data for Image Classification

### 2015-04-14 paper.tex
Final version for camera ready

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\linewidth]{figure/pgm.pdf}
\end{center}
\caption{Probabilistic graphical model of label noise}
\label{fig:noise_pgm}
\end{figure}

@@

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$.
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$.
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$.
\end{enumerate}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/diagram.pdf}
\end{center}
\caption{System diagram of our method. Two CNNs are used to predict the class label $p(\vec{y}|\vec{x})$ and the noise type $p(\vec{z}|\vec{x})$, respectively. The label noise model layer uses both these predictions and the given noisy label to estimate a posterior distribution of the true label, which is then used to supervise the training of CNNs. Data with clean labels are also mixed in to prevent the models from drifting away.}
\label{fig:framework}
\end{figure*}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/cm_strong_weak_labels.pdf}
\end{center}
\caption{Confusion matrix between clean and noisy labels. We hide extremely small grid numbers for better demonstration. Frequency of each true label is listed at the top of each column. The overall accuracy is $61.54\%$, which indicates that the noisy labels are not reliable.}
\label{fig:cm_strong_weak}
\end{figure}
% subsection dataset (end)

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{c|c|c|c|c}
\hline
\# & Method & Training Data & Initialization & Test Accuracy\\
\hline\hline
1 & AlexNet & $\mathcal{D}_c$ & random & 64.54\% \\
2 & AlexNet & $\mathcal{D}_c$ & ImageNet pretrained & 72.63\% \\
3 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & random & 74.03\% \\
4 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & ImaegNet pretrained & 75.13\% \\
5 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & model \#2 & 75.30\% \\
6 & Pseudo-Label\cite{lee2013pseudo} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as unlabeled & model \#2 & 73.04\% \\
7 & Bottom-Up~\cite{sukhbaatar2014learning} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & 76.22\% \\
8 & Ours & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & \textbf{78.24\%} \\
\hline
\end{tabular}
\end{center}
\caption{Experiment results on the collected dataset}
\label{tab:exp}
\end{table*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/model_output.pdf}
\caption{Examples of handling noisy labels. The information layout in each block is illustrated in the top-left one.}
\label{fig:model_output}
\end{center}
\end{figure*}

@@

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|c}
\hline
Noise Level & Base Model & \cite{sukhbaatar2014learning} & Ours \\
\hline\hline
30\% & 65.57\% & 69.73\% & 69.81\% \\
\hline
40\% & 62.38\% & 66.66\% & 66.76\% \\
\hline
50\% & 57.36\% & 63.39\% & 63.00\% \\
\hline
\end{tabular}
\end{center}
\caption{Accuracies on CIFAR-10 with synthetic label noise}
\label{tab:accuracy_synthetic}
\end{table}

@@

Next we demonstrate the effectiveness of learning to predict the label noise type. Notice that if an image has low probability of ``noise free'' (\ie, $p(\vec{z}_1=1|\vec{x})$ is small), then our model will believe it is likely to be mislabeled. In order to check the reliability of these predictions, we first estimate $p(\vec{z}_1=1|\vec{x})$ on the validation set and sort the images accordingly in ascending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that our model can identify mislabeled samples quite well based on their image content, which verifies the observation that mislabeled images often share similar visual patterns.

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_prediction_rank_precision.pdf}
\end{center}
\caption{Rank-precision curve of label noise predictions}
\label{fig:noise_prediction_rp}
\end{figure}
% subsection effect_of_noise_estimation (end) 
% section experiments (end)

>>>>

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\linewidth]{figure/pgm.pdf}
\end{center}
\caption{Probabilistic graphical model of label noise}
\label{fig:noise_pgm}
\end{figure}

@@

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$.
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$.
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$.
\end{enumerate}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.85\linewidth]{figure/diagram.pdf}
\end{center}
\caption{System diagram of our method. Two CNNs are used to predict the class label $p(\vec{y}|\vec{x})$ and the noise type $p(\vec{z}|\vec{x})$, respectively. The label noise model layer uses both these predictions and the given noisy label to estimate a posterior distribution of the true label, which is then used to supervise the training of CNNs. Data with clean labels are also mixed in to prevent the models from drifting away.}
\label{fig:framework}
\end{figure*}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.9\linewidth]{figure/cm_strong_weak_labels.pdf}
\end{center}
\caption{Confusion matrix between clean and noisy labels. We hide extremely small grid numbers for better demonstration. Frequency of each true label is listed at the top of each column. The overall accuracy is $61.54\%$, which indicates that the noisy labels are not reliable.}
\label{fig:cm_strong_weak}
\end{figure}
% subsection dataset (end)

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{c|c|c|c|c}
\hline
\# & Method & Training Data & Initialization & Test Accuracy\\
\hline\hline
1 & AlexNet & $\mathcal{D}_c$ & random & 64.54\% \\
2 & AlexNet & $\mathcal{D}_c$ & ImageNet pretrained & 72.63\% \\
3 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & random & 74.03\% \\
4 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & ImaegNet pretrained & 75.13\% \\
5 & AlexNet & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as ground truths & model \#2 & 75.30\% \\
6 & Pseudo-Label\cite{lee2013pseudo} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ as unlabeled & model \#2 & 73.04\% \\
7 & Bottom-Up~\cite{sukhbaatar2014learning} & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & 76.22\% \\
8 & Ours & upsampled $\mathcal{D}_c$ and $\mathcal{D}_{\eta}$ & model \#2 & \textbf{78.24\%} \\
\hline
\end{tabular}
\end{center}
\caption{Experimental results on the clothing classification dataset. $\mathcal{D}_c$ contains $47,570$ clean labels while $\mathcal{D}_{\eta}$ contains $10^6$ noisy labels.}
\label{tab:exp}
\end{table*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\linewidth]{figure/model_output.pdf}
\caption{Examples of handling noisy labels. The information layout for each block is illustrated on the top-left. $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ are predictions of the true label and noise type based on image content. After observing the noisy label, our model infers the posterior distributions $p(\vec{y}|\vec{\tilde{y}},\vec{x})$ and $p(\vec{z}|\vec{\tilde{y}},\vec{x})$, then replaces the $\vec{y}$ and $\vec{z}$ with them as supervisions to the CNNs.}
\label{fig:model_output}
\end{center}
\end{figure*}

@@

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|c}
\hline
Noise Level & CIFAR10-quick & \cite{sukhbaatar2014learning} & Ours \\
\hline\hline
30\% & 65.57\% & 69.73\% & 69.81\% \\
\hline
40\% & 62.38\% & 66.66\% & 66.76\% \\
\hline
50\% & 57.36\% & 63.39\% & 63.00\% \\
\hline
\end{tabular}
\end{center}
\caption{Accuracies on CIFAR-10 with synthetic label noises. The label noises generated here only depend on the true labels but not the image content, which exactly match the assumption of~\cite{sukhbaatar2014learning} but are unfavored to our model.}
\label{tab:accuracy_synthetic}
\end{table}

@@

Next we demonstrate the effectiveness of learning to predict the label noise type. Notice that if an image has low probability of ``noise free'' (\ie, $p(\vec{z}_1=1|\vec{x})$ is small), then our model will believe it is likely to be mislabeled. In order to check the reliability of these predictions, we estimate $p(\vec{z}_1=1|\vec{x})$ on the validation set and sort the images accordingly in ascending order. For the precision calculation, we consider a candidate image as true positive if its clean label mismatches its original noisy label, and our model predicts it as not ``noisy free''. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that our model can identify mislabeled samples quite well based on their image content, which verifies the observation that mislabeled images often share similar visual patterns.

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_prediction_rank_precision.pdf}
\end{center}
\caption{Rank-precision curve of label noise predictions. We rank the validation images from low to high according to their ``noise free'' probabilities. For the precision calculation, we consider a candidate image as true positive if its clean label mismatches its original noisy label, and our model predicts it as not ``noisy free''.}
\label{fig:noise_prediction_rp}
\end{figure}
% subsection effect_of_noise_estimation (end) 
% section experiments (end)
